Compositional equivalences based on Open pNets

### 2020-04-17 2016-Weak-FH-Bisim/WeakBisim.tex
in and out made visible...
J'ai change les \underline{in|out} partout dans le texte.
Rabea, pourra-tu le faire dans les figures ? (c'ets dans toutes sauf la 7 ....)

@@

Fig. \ref{SimpleProt:Spec} (left) shows the example principle, which corresponds to the hierarchical structure of a pNet: two unspecified
processes $P$ and $Q$ (holes) communicate messages, with a data value
argument, through the two protocol entities. Process $P$ sends an $\texttt{p-send(m)}$ message to the $\symb{Sender}$; this communication is denoted as $\texttt{\underline{in(m)}}$, where underline actions represent internal actions that cannot be further synchronised with the environment, similarly to the $\tau$ of CCS. In our formalisation we will also denote these actions $\tau$ for simplicity of notations but the examples are easier to read when invisible actions have significant names.
At the other end, process $Q$ receives the message from the $\symb{Receiver}$. The holes $P$ and $Q$ can also have other interactions with their environment, represented here by actions $\texttt{p-a}$ and $\texttt{q-b}$. The underlying network is modelled by a medium entity transporting messages from the sender to the receiver, and that is able to detect transport errors and signal them to the sender. The return $\symb{ack}$ message from $\symb{Receiver}$ to $\symb{Sender}$ is supposed to be safe. The final transmission of the message to the recipient (the hole Q) includes the value of the ``error counter'' $ec$.

Fig. \ref{SimpleProt:Spec} (right) shows a graphical view of the pNet that specifies the system:
the full system should be equivalent (e.g. through weak bisimulation)
to the two processes connected simply through a perfect medium.  
The pNet has a tree-like structure.  The root node of the tree {\it SimpleSystemSpec} is the top level of the pNet structure. It acts as the parallel operator. It consists of three nodes: two Holes $P$ and $Q$ and one subnet, denoted {\it PrefectBuffer}. Nodes of the tree are synchronized using four synchronization vectors, that  express the possible synchronizations between the parameterized actions of a subset of the nodes. For instance, in the vector "$< \texttt{p-send(m)},\texttt{in(m)},\_> \rightarrow \texttt{\underline{in(m)}}$ only $P$ and {\it PrefectBuffer} nodes are involved in the synchronization. The synchronization between these processes occurs when  process $P$  performs $\texttt{p-send(m)}$  action sending a message, and  the {\it PrefectBuffer} accepts the message through $\texttt{in(m)}$ action  at the same time; the result that will be returned at upper level is the internal action  $\texttt{\underline{in(m)}}$.

@@

 \smallskip\noindent
 \[  OT_1  = \openrule{
      \{\texttt{P}\mapsto \texttt{p-send(m)}\},  [\texttt{m=m'}],
        \{\texttt{s1\_msg}\gets \texttt{m}\}
                      }
    {\ostate{s_0,m_0,r_0} \OTarrow{\underline{\texttt{in(m')}}} \ostate{s_1,m_0,r_0}}
    \]

@@

 %% \smallskip\noindent
 %% $  OT_2  = \openrule{
 %%   \emptyset; Pred=[send(SV2\_m)=send(s_1\_m);in(SV2\_m)=in(?m)];  \\ Post=\{m_1\_m:=?m\}
 %%                      }
 %%    {\ostate{s_1,m_0,r_0} \OTarrow{\underline{send(SV2\_m)}} \ostate{s_2,m_1,r_0}}
 %%    $

@@

 %%     $  OT_2  = \openrule{
 %%      \emptyset; Pred=\True;  \\ Post=\{m_1\_m:=s_1\_m\}
 %%                      }
 %%    {\ostate{s_1,m_0,r_0} \OTarrow{\underline{send(s_1\_m)}} \ostate{s_2,m_1,r_0}}
 %%    $

@@

%% \LUDO{Je sais pas de quoi parle le prochain todo :-( l'exemple?}
%% \TODO{Ceci n'est pas nouveau, n'est-ce pas, est-ce utile pour le lecteur ici ? Ou est-ce qu'on bouge tous ces ``details techniques'' dans une annexe ? Si oui, je refais avec un morceau de ABP}
%% \begin{example} \emph{Using the operational rules to compute
%%     open-transitions}
%%   In Fig. \ref{usingrules:OT2} we show the deduction tree used to construct and prove the 
%%   open transition $OT_2$ of \texttt{EnableCompL} (see example page \pageref{OT:ABP-composed}).
%%   The rule uses \TrUn\ for the $\delta$ transition of $C_3$, for the $l$ transition of $C_4$, then combines the result using the $a_4$ vector of the bottom pNet node, and the $\underline{\delta(x)}$ vector of the top node.

@@

\begin{example}
  Figure \ref{WOT2} shows the construction of one of the weak transitions of the Specification OA. On the top we show the subset of the original open automaton (from Figure \ref{SimpleProtCounter:SpecOA}) considered here, and at the bottom the generated weak transition.  For reasons of readability, we abbreviate the weak open transitions encoded by $\openrule   {\{\}, True,	() } {s \OTWeakarrow \tau s'}$  as $W_\tau$. The weak open transition shown here is the transition delivering the result of the algorithm to hole $Q$ by applying rules: \WTUn,\WTDeux, and \WTTrois. First rule \WTUn~ adds an $WT_\tau$ loop on each state. Rule \WTDeux~ transforms each 3 OTs into WOTs.   Then consider application of Rule \WTTrois~ on a sequence 3  WOTs.   $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {\text{b1}\OTWeakarrow \tau \text{b1}}$; $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$;  $\openrule
         {\{\}, True,
			() }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$. Then Rule \WTTrois~ produces  $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+2) }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$. We can  can iterate this construction an arbitrary number of times, getting for any natural number $n$ the weak open transition:
  $\openrule
         {\emptyset, True,
			(\texttt{ec}\gets \texttt{ec}+n) }
         {\text{b1} \OTWeakarrow \tau \text{b1}} \forall n \geq 0$.  Finally,  applying again \WTTrois, and using the central open transition having \texttt{\underline{out(m,ec)}}  as $\alpha$, we get the resulting weak open transition between b1 and b0 (as shown in Figure \ref{WOT2}).
 %     $\openrule
 %        {\{\texttt{Q}\mapsto\texttt{q-revc(ec)}\subst{ec\gets ec+n}\}, True,
%			(\texttt{ec}\gets \texttt{ec}+n) }
 %        {b1 \OTWeakarrow {\underline{\texttt{out(m,ec)}}\subst{ec\gets ec+n}} b0} \forall n \geq 0$. 
\end{example}

@@

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
        $$\forall n\ge 0$

@@

$ WI_{8c}(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(s1\_msg,s2\_ec}+n)\}, True, ()}
         {210 \OTWeakarrow {\underline{\texttt{out(r1\_msg,r1\_ec)}\subst{(\texttt{r1\_msg}\gets \texttt{s1\_msg}, \texttt{r1\_ec}\gets \texttt{s2\_msg}+n)}}} 202}
        $($\forall n\ge 0$)

@@

  $ SS_3 = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m1)}\}, True,
    (\texttt{b1\_msg}\gets \texttt{m1}, \texttt{b1\_ec}\gets 0)}
  {\text{b0} \OTarrow {\underline{\texttt{in(m1)}}} \text{b1}}
  $

$ WI_3(0) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m2)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m2}, \texttt{s1\_ec}\gets 0)}
  {000 \OTWeakarrow {\underline{\texttt{in(m2)}}} 100}
  $

@@

$\forall \texttt{m1}. \{True \land True\hspace{3mm} \implies
\exists \texttt{m2}. \\
(\texttt{[p-send(m1)]} = \texttt{[p-send(m2)]}
\land True
\land \underline{\texttt{in(m1)}} = \underline{\texttt{in(m2)}} \land 
\\
(\texttt{b1\_msg\!\! = \!\!s1\_msg} \land \texttt{b1\_ec\!\! =\!\! s1\_ec}) \subst{(\texttt{b1\_msg}\!\gets\! \texttt{m1}, \texttt{b1\_ec}\!\gets\! 0)\uplus (\texttt{s1\_msg}\!\gets\! \texttt{m2},\texttt{s1\_ec}\!\gets\! 0)})\}$

@@

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
  { \{000,202\} \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
$

$ WI_{3a}(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{m1\_msg}\gets \texttt{m}, \texttt{m1\_ec}\gets n, \texttt{s2\_ec}\gets n)}
  { \{000,202\} \OTWeakarrow {\underline{\texttt{in(m)}}} 210}
$

$ WI_{3b}(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s2\_ec}\gets n)}
  { \{000,202\} \OTWeakarrow {\underline{\texttt{in(m)}}} 220}
$

$ WI_{3c}(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{r1\_msg}\gets \texttt{m}, \texttt{r1\_ec}\gets n)}
  { \{000,202\} \OTWeakarrow {\underline{\texttt{in(m)}}} 201}
$

@@

$ WI_8 = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(r1-msg,r1-ec)}\}, True, ()}
         {201 \OTWeakarrow {\underline{\texttt{out(r1-msg,r1-ec)}}} \{202, 000\}}
         \ \forall n\ge 0$

$ WI_{8a}(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(m1\_msg,m1\_ec}+n)\}, True, ()}
         {210 \OTWeakarrow {\underline{\texttt{out(m1\_msg,m1\_ec}+n)}} \{202, 000\}}
         \ \forall n\ge 1$

 $ WI_{8b}(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(??\_msg,s2\_ec}+n)\}, True, ()}
         {220 \OTWeakarrow {\underline{\texttt{out(??\_msg,m2\_ec}+n)}} \{202, 000\}}
         \ \forall n\ge 1$

 $ WI_{8c}(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(s1\_msg,s1\_ec}+n)\}, True, ()}
         {100 \OTWeakarrow {\underline{\texttt{out(s1\_msg,s1\_ec}+n)}} \{202, 000\}}
         \ \forall n\ge 1$

>>>>

\newcommand{\nounderline}[1]{#1}

@@

\TODO{J'ai change les $\underline{in}$ et $\underline{out}$ en $in$ et $out$ dans le texte. Rabea, peux-tu le faire dans toutes les figures (1, 2, 3,4,5,6) ?}

Fig. \ref{SimpleProt:Spec} (left) shows the example principle, which corresponds to the hierarchical structure of a pNet: two unspecified
processes $P$ and $Q$ (holes) communicate messages, with a data value
argument, through the two protocol entities. Process $P$ sends an $\texttt{p-send(m)}$ message to the $\symb{Sender}$; this communication is denoted as $\texttt{\nounderline{in(m)}}$% , where underline actions represent internal actions that cannot be further synchronised with the environment, similarly to the $\tau$ of CCS. In our formalisation we will also denote these actions $\tau$ for simplicity of notations but the examples are easier to read when invisible actions have significant names}
.
At the other end, process $Q$ receives the message from the $\symb{Receiver}$. The holes $P$ and $Q$ can also have other interactions with their environment, represented here by actions $\texttt{p-a}$ and $\texttt{q-b}$. The underlying network is modelled by a medium entity transporting messages from the sender to the receiver, and that is able to detect transport errors and signal them to the sender. The return $\symb{ack}$ message from $\symb{Receiver}$ to $\symb{Sender}$ is supposed to be safe. The final transmission of the message to the recipient (the hole Q) includes the value of the ``error counter'' $ec$.

Fig. \ref{SimpleProt:Spec} (right) shows a graphical view of the pNet that specifies the system:
the full system should be equivalent (e.g. through weak bisimulation)
to the two processes connected simply through a perfect medium.  
The pNet has a tree-like structure.  The root node of the tree {\it SimpleSystemSpec} is the top level of the pNet structure. It acts as the parallel operator. It consists of three nodes: two Holes $P$ and $Q$ and one subnet, denoted {\it PrefectBuffer}. Nodes of the tree are synchronized using four synchronization vectors, that  express the possible synchronizations between the parameterized actions of a subset of the nodes. For instance, in the vector "$< \texttt{p-send(m)},\texttt{in(m)},\_> \rightarrow \texttt{\nounderline{in(m)}}$" only $P$ and {\it PrefectBuffer} nodes are involved in the synchronization. The synchronization between these processes occurs when  process $P$  performs $\texttt{p-send(m)}$  action sending a message, and  the {\it PrefectBuffer} accepts the message through $\texttt{in(m)}$ action  at the same time; the result that will be returned at upper level is the action  $\texttt{\nounderline{in(m)}}$.

@@

 \smallskip\noindent
 \[  OT_1  = \openrule{
      \{\texttt{P}\mapsto \texttt{p-send(m)}\},  [\texttt{m=m'}],
        \{\texttt{s1\_msg}\gets \texttt{m}\}
                      }
    {\ostate{s_0,m_0,r_0} \OTarrow{\nounderline{\texttt{in(m')}}} \ostate{s_1,m_0,r_0}}
    \]

@@

 %% \smallskip\noindent
 %% $  OT_2  = \openrule{
 %%   \emptyset; Pred=[send(SV2\_m)=send(s_1\_m);in(SV2\_m)=in(?m)];  \\ Post=\{m_1\_m:=?m\}
 %%                      }
 %%    {\ostate{s_1,m_0,r_0} \OTarrow{\nounderline{send(SV2\_m)}} \ostate{s_2,m_1,r_0}}
 %%    $

@@

 %%     $  OT_2  = \openrule{
 %%      \emptyset; Pred=\True;  \\ Post=\{m_1\_m:=s_1\_m\}
 %%                      }
 %%    {\ostate{s_1,m_0,r_0} \OTarrow{\nounderline{send(s_1\_m)}} \ostate{s_2,m_1,r_0}}
 %%    $

@@

%% \LUDO{Je sais pas de quoi parle le prochain todo :-( l'exemple?}
%% \TODO{Ceci n'est pas nouveau, n'est-ce pas, est-ce utile pour le lecteur ici ? Ou est-ce qu'on bouge tous ces ``details techniques'' dans une annexe ? Si oui, je refais avec un morceau de ABP}
%% \begin{example} \emph{Using the operational rules to compute
%%     open-transitions}
%%   In Fig. \ref{usingrules:OT2} we show the deduction tree used to construct and prove the 
%%   open transition $OT_2$ of \texttt{EnableCompL} (see example page \pageref{OT:ABP-composed}).
%%   The rule uses \TrUn\ for the $\delta$ transition of $C_3$, for the $l$ transition of $C_4$, then combines the result using the $a_4$ vector of the bottom pNet node, and the $\nounderline{\delta(x)}$ vector of the top node.

@@

\begin{example}
  Figure \ref{WOT2} shows the construction of one of the weak transitions of the Specification OA. On the top we show the subset of the original open automaton (from Figure \ref{SimpleProtCounter:SpecOA}) considered here, and at the bottom the generated weak transition.  For reasons of readability, we abbreviate the weak open transitions encoded by $\openrule   {\{\}, True,	() } {s \OTWeakarrow \tau s'}$  as $W_\tau$. The weak open transition shown here is the transition delivering the result of the algorithm to hole $Q$ by applying rules: \WTUn,\WTDeux, and \WTTrois. First rule \WTUn~ adds an $WT_\tau$ loop on each state. Rule \WTDeux~ transforms each 3 OTs into WOTs.   Then consider application of Rule \WTTrois~ on a sequence 3  WOTs.   $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {\text{b1}\OTWeakarrow \tau \text{b1}}$; $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+1) }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$;  $\openrule
         {\{\}, True,
			() }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$. Then Rule \WTTrois~ produces  $\openrule
         {\{\}, True,
			(\texttt{ec}\gets \texttt{ec}+2) }
         {\text{b1} \OTWeakarrow \tau \text{b1}}$. We can  can iterate this construction an arbitrary number of times, getting for any natural number $n$ the weak open transition:
  $\openrule
         {\emptyset, True,
			(\texttt{ec}\gets \texttt{ec}+n) }
         {\text{b1} \OTWeakarrow \tau \text{b1}} \forall n \geq 0$.  Finally,  applying again \WTTrois, and using the central open transition having \texttt{\nounderline{out(m,ec)}}  as $\alpha$, we get the resulting weak open transition between b1 and b0 (as shown in Figure \ref{WOT2}).
 %     $\openrule
 %        {\{\texttt{Q}\mapsto\texttt{q-revc(ec)}\subst{ec\gets ec+n}\}, True,
%			(\texttt{ec}\gets \texttt{ec}+n) }
 %        {b1 \OTWeakarrow {\nounderline{\texttt{out(m,ec)}}\subst{ec\gets ec+n}} b0} \forall n \geq 0$. 
\end{example}

@@

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
  {000 \OTWeakarrow {\nounderline{\texttt{in(m)}}} 100}
        $$\forall n\ge 0$

@@

$ WI_{8c}(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(s1\_msg,s2\_ec}+n)\}, True, ()}
         {210 \OTWeakarrow {\nounderline{\texttt{out(r1\_msg,r1\_ec)}\subst{(\texttt{r1\_msg}\gets \texttt{s1\_msg}, \texttt{r1\_ec}\gets \texttt{s2\_msg}+n)}}} 202}
        $($\forall n\ge 0$)

@@

  $ SS_3 = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m1)}\}, True,
    (\texttt{b1\_msg}\gets \texttt{m1}, \texttt{b1\_ec}\gets 0)}
  {\text{b0} \OTarrow {\nounderline{\texttt{in(m1)}}} \text{b1}}
  $

$ WI_3(0) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m2)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m2}, \texttt{s1\_ec}\gets 0)}
  {000 \OTWeakarrow {\nounderline{\texttt{in(m2)}}} 100}
  $

@@

$\forall \texttt{m1}. \{True \land True\hspace{3mm} \implies
\exists \texttt{m2}. \\
(\texttt{[p-send(m1)]} = \texttt{[p-send(m2)]}
\land True
\land \nounderline{\texttt{in(m1)}} = \nounderline{\texttt{in(m2)}} \land 
\\
(\texttt{b1\_msg\!\! = \!\!s1\_msg} \land \texttt{b1\_ec\!\! =\!\! s1\_ec}) \subst{(\texttt{b1\_msg}\!\gets\! \texttt{m1}, \texttt{b1\_ec}\!\gets\! 0)\uplus (\texttt{s1\_msg}\!\gets\! \texttt{m2},\texttt{s1\_ec}\!\gets\! 0)})\}$

@@

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
  { \{000,202\} \OTWeakarrow {\nounderline{\texttt{in(m)}}} 100}
$

$ WI_{3a}(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{m1\_msg}\gets \texttt{m}, \texttt{m1\_ec}\gets n, \texttt{s2\_ec}\gets n)}
  { \{000,202\} \OTWeakarrow {\nounderline{\texttt{in(m)}}} 210}
$

$ WI_{3b}(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s2\_ec}\gets n)}
  { \{000,202\} \OTWeakarrow {\nounderline{\texttt{in(m)}}} 220}
$

$ WI_{3c}(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{r1\_msg}\gets \texttt{m}, \texttt{r1\_ec}\gets n)}
  { \{000,202\} \OTWeakarrow {\nounderline{\texttt{in(m)}}} 201}
$

@@

$ WI_8 = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(r1-msg,r1-ec)}\}, True, ()}
         {201 \OTWeakarrow {\nounderline{\texttt{out(r1-msg,r1-ec)}}} \{202, 000\}}
         \ \forall n\ge 0$

$ WI_{8a}(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(m1\_msg,m1\_ec}+n)\}, True, ()}
         {210 \OTWeakarrow {\nounderline{\texttt{out(m1\_msg,m1\_ec}+n)}} \{202, 000\}}
         \ \forall n\ge 1$

 $ WI_{8b}(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(??\_msg,s2\_ec}+n)\}, True, ()}
         {220 \OTWeakarrow {\nounderline{\texttt{out(??\_msg,m2\_ec}+n)}} \{202, 000\}}
         \ \forall n\ge 1$

 $ WI_{8c}(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(s1\_msg,s1\_ec}+n)\}, True, ()}
         {100 \OTWeakarrow {\nounderline{\texttt{out(s1\_msg,s1\_ec}+n)}} \{202, 000\}}
         \ \forall n\ge 1$
